---
title: "HW10"
date: "due 4/10/2019"
output:
  pdf_document: default
  html_document: default
---
###Instructions. We will again use a SEER breast cancer dataset (BRCA_SEER_SURV.xlsx on Canvas listed as SEER dataset for HW9) that includes first primary malignancy breast cancer cases diagnosed from 2007-2009 who were in the SEER 18 database for this homework assignment. The code for loading the dataset is provided below along with data management code for the variables that will be used to solve the problems. Please submit your homework by uploading the .RMD file or the HTML NB file to Blackboard under the HW9 assignment.

###Import the data and library packages
```{r, warning=FALSE, eval=FALSE}
#install.packages("survminer")#for pairwise diffs
library(readr) #for read txt file
library(survival) #for calculating KM values
library(survminer)#for pairwise diffs
library(ggfortify) #for KM curves
library(readxl) # for reading in excel file
library(ggplot2) # for plotting KM curve
library(tidyverse) # for various packages
library(lmtest) #model comparison
library(stargazer) #models presentation
BRCA <- read_excel("BRCA_SEER_SURV.xlsx", sheet=1) #load data for this exercise
str(BRCA)
```
###Data management for variables used in this problem set (stage_f, event_f, and insurance_f)
```{r, warning=FALSE, eval=FALSE}
#provide shorter names for variables
names(BRCA)<-c("ID", "age_dx","yr_dx", "sex", "race", "ishispanic", 
               "insurance", "marital", "pov_pct", "edu_pct", "cause_spec_death", 
               "cause_other_death", "surv_mo", "vital_stat", "Stage") 

##Recode stage variable as a factor variable and label it
BRCA$stage_f[
    BRCA$Stage=="I"]<-0
BRCA$stage_f[
    BRCA$Stage=="IIA"|
    BRCA$Stage=="IIB"]<-1
BRCA$stage_f[
    BRCA$Stage=="IIIA"|
    BRCA$Stage=="IIIB"|
    BRCA$Stage=="IIIC"|
    BRCA$Stage=="IIINOS"]<-2
BRCA$stage_f[
    BRCA$Stage=="IV"]<-3
BRCA$stage_f[
    BRCA$Stage=="UNK Stage"]<-4
BRCA$stage_f<-factor(BRCA$stage_f,
      levels = c(0,1,2,3,4),
      labels = c("Stage 1", "Stage 2", "Stage 3", "Stage 4", "Stage Unknown"))

#Recode cause specific death as 1/0 if the person died/did not die of breast cancer 
BRCA$event_f[
    BRCA$cause_spec_death=="Dead (attributable to this cancer dx)"]<-1
BRCA$event_f[
    BRCA$cause_spec_death=="Alive or dead of other cause"|
    BRCA$cause_spec_death=="N/A not first tumor"]<-0

#Recode insurance status as a factor variable and label it
BRCA$insurance_f[
  BRCA$insurance=="Insured"]<-0
BRCA$insurance_f[
  BRCA$insurance=="Insured/No specifics"]<-1
BRCA$insurance_f[
  BRCA$insurance=="Any Medicaid"]<-2
BRCA$insurance_f[
  BRCA$insurance=="Uninsured"]<-3
BRCA$insurance_f[
  BRCA$insurance=="Insurance status unknown"]<-4
BRCA$insurance_f<-factor(BRCA$insurance_f,
      levels = c(0,1,2,3,4),
      labels = c("Insured", "Insured/No specifics", "Any Medicaid", "Uninsured", "Unknown"))
```
